---
title: Calling R from Clojure with rojure
author: Carsten Behring
date: '2017-10-04'
slug: calling-r-from-clojure-with-rojure
categories: []
tags:
  - clojure
  - R 
---

```{r, echo=F}
knitr::opts_chunk$set(cache=T)

``` 
In this short post I will show, how we can call R from Clojure, by using the [rojure](https://github.com/behrica/rojure) package.


# Requirements 

At first we need to add rojure to the dependencies of the Clojure project, we work on:

``` clojure
[rojure "0.2.0"]
```

Secondly we need to start an R process on our machine, and install(if not done)  and use the "Rserve" package.


``` R
# install.packages("Rserve")
library(Rserve)
Rserve()

```
This will start an server process, which the rojure library can connect to and send it R data and code for execution.

# Examples

## Evaluating R code

This is the simplest usage.
We can evaluate arbitrary R code, given as a string:

```{r engine="lein"}
(use '(rojure core))
(use '[clojure.core.matrix :as m])
(def r (get-r))
(println (r-eval r "sessionInfo()"))
```

## Clojure -> R

Clojure objects, can be set in the R session. This works 
for:

* sequences / vectors
* core.matrix dataset
* core.matrix matrix



## R->Clojure 

Getting any data.frame from R returns it as object of type core.matrix dataset.


```{r engine="lein"}
(use '(rojure core))
(use '[clojure.core.matrix :as m]
     '[clojure.core.matrix.dataset :as ds]

)
(def r (get-r))
(def iris (r-get r "iris"))
(println "shape: " (m/shape iris))
(println "column names: " (ds/column-names iris))
(println "class: " (class iris))
(println "First 5 of Sepal.Length" (take 5 (ds/column iris "Sepal.Length")))
```


The same works for matrices, which becomes a data structure on which the functions from 
clojre.core.matrix can be called.

```{r engine="lein"}
(use '(rojure core))
(use '[clojure.core.matrix :as m])
(def r (get-r))
(def m (r-eval r "diag(1, 4, 4)"))
(println "shape: "(m/shape m))
(println "class " (class m))
(println (m/pm m))
```

The concrete implementation of the matrix class created, depends on the current default,
which can be changed by calling "m/set-current-implementation"


# Using r-transform to train/predict a classifier 

In the following example I call the quick start code from R  package [mlr](https://mlr-org.github.io/mlr-tutorial/release/html/) from Clojure for training a classifier.

The "iris" data comes from "incanter/get-dataset", gets written into the R session, 
and then we call the mlr R functions from Clojure.

I demonstrate two version of the same R code, by using either "with-r-eval" "r-transform"

## with with-r-eval

Here the R code is given as sequence of strings.


```{r engine="lein"}
(use '(rojure core))
(use '[clojure.core.matrix :as m])
(require '[incanter.core :as ic])
(require '[incanter.datasets :as id])

(def r (get-r))
(def iris-ds (id/get-dataset :iris))
; change col names, mlr package des not like ":" in variables     	  
(def iris-ds (ic/rename-cols {:Sepal.Length "Sepal.Length"
                             :Sepal.Width "Sepal.Width"
                             :Petal.Length "Petal.Length"
                             :Petal.Width "Petal.Width"
                             :Species "Species"
                              } iris-ds))

(r-set! r "iris" iris-ds)

 (println (with-r-eval r
  "library(mlr)"
  "task = makeClassifTask(data = iris, target = 'Species')"
  "lrn = makeLearner('classif.lda')"
  "n = nrow(iris)"
  "train.set = sample(n, size = 2/3*n)"
  "test.set = setdiff(1:n, train.set)"
  "model = train(lrn, task, subset = train.set)"
  "pred = predict(model, task = task, subset = test.set)"
  "performance(pred, measures = list(mmce, acc))"
  ))


```

## Using r-transform

The same can be achieved by calling r-transform and giving it a R script to execute.



```{r engine="lein"}
(use '(rojure core))
(use '[clojure.core.matrix :as m])
(require '[incanter.core :as ic])
(require '[incanter.datasets :as id])

(def r (get-r))
(def iris-ds (id/get-dataset :iris))
; change col names, mlr package des not like ":" in variables     	  
(def iris-ds (ic/rename-cols {:Sepal.Length "Sepal.Length"
                             :Sepal.Width "Sepal.Width"
                             :Petal.Length "Petal.Length"
                             :Petal.Width "Petal.Width"
                             :Species "Species"
                              } iris-ds))

		
(println (r-transform r iris-ds "mlr.R"))
``` 

The R code in 'mlr.R' code looks likes this:
(Notice the usage of 'in_' and 'Out_' for input and output data.frames.)


``` R
library(mlr)
task = makeClassifTask(data = in_, target = "Species")
lrn = makeLearner("classif.lda")
n = nrow(in_)
train.set = sample(n, size = 2/3*n)
test.set = setdiff(1:n, train.set)
model = train(lrn, task, subset = train.set)
pred = predict(model, task = task, subset = test.set)
out_ <- performance(pred, measures = list(mmce, acc))
```


## Transforming of S3 classes


R functions which return S3 objects gets returned to Clojure as maps.

```{r engine="lein"}
(use '(rojure core))
(use '[clojure.core.matrix :as m])
(def r (get-r))
(println (r-eval r "sessionInfo()"))
```

## Clojure -> R

Clojure objects, can be set in the R session. This works 
for:

* sequences / vectors
* core.matrix dataset
* core.matrix matrix


## R->Clojure 

Getting any data.frame from R returns it as object of type core.matrix dataset.

```{r engine="lein"}
(use '(rojure core))
(use '[clojure.core.matrix :as m])
(def r (get-r))
(def iris (r-get r "iris"))
(println "shape: " (m/shape iris))
(println "class: " (class iris))
```


The same works for matrices, which becomes a data structure on which the functions from 
clojre.core.matrix can be called.

S3 classes gets converted to maps-of-maps, for example the class "lm" produced by the R function "lm".

```{r engine="lein"}
(use '(rojure core))
(def r (get-r))
(println (keys (r-eval r "lm(Fertility ~ . , data = swiss)")))

(println (get (r-eval r "lm(Fertility ~ . , data = swiss)")
     "coefficients"))
```


# Anecdote

This blogs is written in R markdown, containing executable code blocks written in Clojure which call again R.....
So R calls Clojure calls R ....
